Adaptive evolution with aneuploidy and mutation

### 2021-09-29 ms/ms.tex
added some todos for ilia
added some figures

\usepackage{amssymb,amsmath,amsthm}
\usepackage{commath}
\usepackage{longtable}
\usepackage[hyphens]{url}
\usepackage[dvipsnames]{xcolor}
\usepackage[unicode=true,colorlinks=true,urlcolor=CadetBlue,citecolor=black,linkcolor=black]{hyperref}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage{authblk}
\usepackage{subcaption}
\captionsetup[subfigure]{labelformat=empty}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{longtable}
\usepackage{siunitx}

@@

\paragraph{Likelihood function.}
Denote by $A_{t}$ the fixation of aneuploidy at generation $t$, by $L_{t}$ the loss of aneuploidy at generation $t$, and by $L^*_{t,T}$ the loss of aneuploidy at generation $t$ conditioned on no loss by generation $T$.
The model likelihood function for parameter vector $\theta$ for the heat-stress experiment is
\begin{multline} \label{eq:heatstress-likelihood}
\likelihood(\theta | A_{450}, L_{1700}, L^*_{2350, 1700} ) = \\
P^4(A_{450}) \cdot \Big(1 - P^4(\lnot{L_{1700}} | A_{450} ) - P^4(\lnot{L^*_{2350, 1700}}| A_{450}) + P^4(\lnot{L_{1700}} \land \lnot{L^*_{2350, 1700}}| A_{450})\Big)	\;,	
\end{multline}
where $P(X)$ is the probability of even $X$ and $\lnot$ means \emph{not}.
The likelihood function of the model with parameter set $\theta$ for the pH-stress experiment is
\begin{equation} \label{eq:phstress-likelihood}
\likelihood(\theta | A_{150}, L_{750} ) = \\
 P^4(L_{450}) \cdot 6 \cdot	P^4(L_{750} | A_{150}) \cdot P^4(\lnot{L_{750}} | A_{150}) \;.
\end{equation} 
Because of the complexity of the model, this likelihood function is intractable. Therefore, given specific values of the model parameters $\theta$, we simulate the model in many replicates to approximate the value of the likelihood function $\likelihood(\theta)$.

@@

>>>>

\usepackage{amssymb,amsmath,amsthm}
\usepackage{commath}
\usepackage{longtable}
\usepackage[hyphens]{url}
\usepackage[dvipsnames]{xcolor}
\usepackage[unicode=true,colorlinks=true,urlcolor=CadetBlue,citecolor=black,linkcolor=black]{hyperref}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage{authblk}
\usepackage{subcaption}
\captionsetup[subfigure]{labelformat=empty}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{longtable}
\usepackage{gensymb,siunitx}

@@

\paragraph{Likelihood function.}
Denote by $A_{t}$ the fixation of aneuploidy at generation $t$, by $L_{t}$ the loss of aneuploidy at generation $t$, and by $L^*_{t,T}$ the loss of aneuploidy at generation $t$ conditioned on no loss by generation $T$.
The model likelihood function for parameter vector $\theta$ for the heat-stress experiment is
\begin{multline} \label{eq:heatstress-likelihood}
\likelihood(\theta | A_{450}, L_{1700}, L^*_{2350, 1700} ) = \\
P^4(A_{450}) \cdot \Big(1 - P^4(\lnot{L_{1700}} | A_{450} ) - P^4(\lnot{L^*_{2350, 1700}}| A_{450}) + P^4(\lnot{L_{1700}} \land \lnot{L^*_{2350, 1700}}| A_{450})\Big)	\;,	
\end{multline}
where $P(X)$ is the probability of event $X$ and $\lnot$ means \emph{not}.
The likelihood function of the model with parameter set $\theta$ for the pH-stress experiment is
\begin{equation} \label{eq:phstress-likelihood}
\likelihood(\theta | A_{150}, L_{750} ) = \\
 P^4(L_{450}) \cdot 6 \cdot	P^4(L_{750} | A_{150}) \cdot P^4(\lnot{L_{750}} | A_{150}) \;.  %%TODO i think it is wrong because fig 5C shows that only 2 replicas are fixated after 650 generations (see Supplementary also)
\end{equation} 
Because of the complexity of the model, this likelihood function is intractable. Therefore, given specific values of the model parameters $\theta$, we simulate the model in many replicates to approximate the value of the likelihood function $\likelihood(\theta)$.

@@

First we evaluated simulations of single-locus model to inference the model parameters. See \ref{fig:posterior} for the posterior. The mutation rate (mean 1.71e-10 (95 CI..)) and trisomy rate (..) corresponds to the literature

\begin{figure}[h]
  \centering
\includegraphics[width=0.9\textwidth]{../figures/posterior.pdf}
  \caption{
  \textbf{Single-locus model parameters posterior distribution}
  The mean and the 95\% confindence interval of the parameters: 
  mutation rate - 1.71e-10 (1.55e-10 - 1.88e-10),
  trisomy rate - 3.3e-05 (2.94e-05 - 3.66e-05),
  2n+1 fitness - 1.05 (1.05 - 1.05),
  2n+1* fitness - 5.24 (4.21 - 6.27),
  2n* fitness - 5.28 (4.24 - 6.32)
  } 
  \label{fig:posterior}
\end{figure}

\begin{figure}[h]
  \centering
\includegraphics[width=0.6\textwidth]{../figures/fixation-plot.pdf}
  \caption{
  \textbf{Time of 2n* fixation can't be explained without aneuploidy}
   Distribution of time to fixation of 2n* and elimination of trisomy for $10,000$ simulations with parameters taken from the mean of the posterior. 
  }
  \label{fig:2n*-fixation}
\end{figure}

\begin{figure}[h]
  \centering
\includegraphics[width=0.6\textwidth]{../figures/ppc-plot.pdf}
  \caption{
  \textbf{Evolutionary dynamics with fitted single-locus model} 
  Frequency of the four genotypes averaged over 10,000 simulations of the model. Each line represent one of the 10 parameter sets drawed from kde of the posterior.
  }
  \label{fig:ppc-plot}
\end{figure}

\begin{figure}[h!]
  \centering
\includegraphics[width=0.6\textwidth]{../figures/likelihood-plot.pdf}
  \caption{
  \textbf{Fit illustration for single locus-model} 
  TODO
  The best fit occurs when first dot is low, last dot is hight, and the inner rectangulars are queal...
  }
  \label{fig:likelihood}
\end{figure}
